Light inextensible strings under tension in the Schwarzschild geometry

### 2020-01-03 string_european_journal_of_physics.tex
use Roman d for differentials as per the style guide

\end{frontmatter}

@@

\begin{equation}\label{schwarzschild}
ds^2= -dt^2\left(1-2M/r\right) +\frac{dr^2}{1-2M/r} + r^2\left(d\theta^2 + \sin^2\theta d\phi^2\right).
\end{equation}

\noindent Here I consider the spatial component~$ds^2=
dr^2/\left(1-2M/r\right) + r^2d\phi^2$, confined to equatorial
points~($\theta=\pi/2$); we understand~$r\geqslant 2M$ throughout.  A
circular string at constant~$r$ will, trivially, have proper
length~$2\pi r$.

@@

\begin{equation}\label{radial_string_length}
  \int_{r=r_1}^{r_2}\frac{dr}{\sqrt{1-2M/r}}=
  \left.
  \sqrt{r(r-2M)} +\frac{1}{2}\log\left(
  \frac{\sqrt{r}+\sqrt{r-2M}}{\sqrt{r}-\sqrt{r-2M}}\right)
  \right|_{r_1}^{r_2}
  \end{equation}

@@

 \begin{equation}\label{ell}
   \ell =
   \int_{r=2M}^{r=2M(1+w)}ds
   =2M\left(2\sqrt{w} + \frac{1}{3}w^{3/2} + O(w^{5/2})\right)
 \end{equation}

@@

\begin{equation}
%  \int_{p_1}^{p_2}\sqrt{\frac{\left(r'\right)^2}{1-2M/r} + r^2}d\phi=
  \int_{p_1}^{p_2}\left(\frac{\left(r'\right)^2}{1-2M/r} + r^2\right)^\frac{1}{2}d\phi=
  \int_{p_1}^{p_2}F\left(r,r'\right)d\phi
\end{equation}

\noindent where dashes denote differentiation with respect to~$\phi$.
The Euler-Lagrange equations for this system,
$\frac{d}{d\phi}\frac{\partial F}{\partial r'}-\frac{\partial
  F}{\partial r}=0$ give an expression for the second derivative
of~$r$ (Mathematica idiom would be {\tt EulerEquations[Sqrt[y[x]\^{}2
      + y'[x]\^{}2/(1-1/y[x])],y[x],x]}):

>>>>

\end{frontmatter}
\newcommand{\rmd}{\mathrm{d}}

@@

\begin{equation}\label{schwarzschild}
\rmd s^2= -\rmd t^2\left(1-2M/r\right) +\frac{\rmd r^2}{1-2M/r} + r^2\left(\rmd\theta^2 + \sin^2\theta\rmd\phi^2\right).
\end{equation}

\noindent Here I consider the spatial component~$\rmd s^2=
\rmd r^2/\left(1-2M/r\right) + r^2\rmd\phi^2$, confined to equatorial
points~($\theta=\pi/2$); we understand~$r\geqslant 2M$ throughout.  A
circular string at constant~$r$ will, trivially, have proper
length~$2\pi r$.

@@

\begin{equation}\label{radial_string_length}
  \int_{r=r_1}^{r_2}\frac{\rmd r}{\sqrt{1-2M/r}}=
  \left.
  \sqrt{r(r-2M)} +\frac{1}{2}\log\left(
  \frac{\sqrt{r}+\sqrt{r-2M}}{\sqrt{r}-\sqrt{r-2M}}\right)
  \right|_{r_1}^{r_2}
  \end{equation}

@@

 \begin{equation}\label{ell}
   \ell =
   \int_{r=2M}^{r=2M(1+w)}\rmd s
   =2M\left(2\sqrt{w} + \frac{1}{3}w^{3/2} + O(w^{5/2})\right)
 \end{equation}

@@

\begin{equation}
%  \int_{p_1}^{p_2}\sqrt{\frac{\left(r'\right)^2}{1-2M/r} + r^2}d\phi=
  \int_{p_1}^{p_2}\left(\frac{\left(r'\right)^2}{1-2M/r} + r^2\right)^\frac{1}{2}\rmd\phi=
  \int_{p_1}^{p_2}F\left(r,r'\right)\rmd\phi
\end{equation}

\noindent where dashes denote differentiation with respect to~$\phi$.
The Euler-Lagrange equations for this system,
$\frac{\rmd}{\rmd\phi}\frac{\partial F}{\partial r'}-\frac{\partial
  F}{\partial r}=0$ give an expression for the second derivative
of~$r$ (Mathematica idiom would be {\tt EulerEquations[Sqrt[y[x]\^{}2
      + y'[x]\^{}2/(1-1/y[x])],y[x],x]}):
